---
title: "Mowzli-Analysis"
output: html_document
date: "2023-12-04"
---

```{r setup, include=FALSE}
# Some of common RMD options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
knitr::opts_chunk$set(warning = F, message = F)
# Can globally set option for number display format.
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
```

```{r include=FALSE}
# Load necessary libraries
library(dplyr)

# Read the dataset
food_access_data <- read.csv("../datasets/food_access_research_atlas.csv")
```

## Median Family Income Analysis

From the previous EDA analysis, we learned that Median Family Income is an important feature contributing to the prevalence of a region being a food desert. To further investigate this, we conducted a t-test over the `MedianFamilyIncome` variable.

```{r include=FALSE}
# Set thresholds for defining potential food deserts
poverty_rate_threshold <- 20  # percent
median_income_threshold <- quantile(food_access_data$MedianFamilyIncome, 0.25)  # lower quartile
snap_beneficiary_threshold <- 20  # percent

# Creating a food desert indicator
food_access_data$PotentialFoodDesert <- with(food_access_data, 
  (PovertyRate >= poverty_rate_threshold) &
  (MedianFamilyIncome <= median_income_threshold) &
  (TractSNAP / POP2010 * 100 >= snap_beneficiary_threshold)
)

# Calculating average median family income for potential food desert areas and non-food desert areas
average_income_food_desert <- mean(food_access_data[food_access_data$PotentialFoodDesert, "MedianFamilyIncome"])
average_income_non_food_desert <- mean(food_access_data[!food_access_data$PotentialFoodDesert, "MedianFamilyIncome"])

# Performing a t-test to check if the difference in median family incomes is statistically significant
t_test_result <- t.test(
  food_access_data[food_access_data$PotentialFoodDesert, "MedianFamilyIncome"],
  food_access_data[!food_access_data$PotentialFoodDesert, "MedianFamilyIncome"]
)

# Print the results
cat("Average Median Family Income in Potential Food Deserts:", average_income_food_desert, "\n")
cat("Average Median Family Income in Non-Food Desert Areas:", average_income_non_food_desert, "\n")
cat("T-Test Results:\n")
print(t_test_result)
```


### T-Test Results

The t-test was conducted to assess the difference in median family incomes between potential food desert areas and non-food desert areas. Here are the results:

Average Median Family Income in Potential Food Deserts: **$22,544.33**
Average Median Family Income in Non-Food Desert Areas: **$67,927.53**

The t-test resulted in a t-statistic of **-149.61** and a p-value of **< 2.2e-16**. The small p-value suggests a significant difference in median family incomes between potential food desert and non-food desert areas.

**Inference:**

Based on the t-test results, we can conclude that there is a statistically significant difference in median family incomes between potential food desert areas and non-food desert areas. Areas classified as potential food deserts tend to have lower median family incomes.


```{r include=FALSE}
logistic_model <- glm(
  PotentialFoodDesert ~ MedianFamilyIncome,
  data = food_access_data,
  family = binomial
)

# Summarize the logistic regression results
summary(logistic_model)

exp(coef(logistic_model))
```

### Logistic Regression Results

A logistic regression model was employed to further investigate the relationship between MedianFamilyIncome and the likelihood of an area being a potential food desert. Here are the results:

The odds ratio for MedianFamilyIncome is approximately **-0.915**. This suggests that for every one-unit increase in `MedianFamilyIncome`, the odds of an area being a potential food desert decrease by a factor of approximately **-0.915**, holding other variables constant.

### Effect of a $1,000 Increase in Median Family Income

To understand the effect of a **$1,000** increase in median family income, we can calculate the change in odds:

Change in Odds = Coefficient for **`MedianFamilyIncome` * 1000**

Change in Odds ≈ **-0.915**

So, a **$1,000** increase in median family income is associated with a decrease in the odds of an area being a potential food desert by a factor of approximately **-0.915**, or a **9.15%** reduction in the odds.

This analysis highlights the importance of `MedianFamilyIncome` in determining the likelihood of an area being classified as a potential food desert. Higher median family incomes are associated with a reduced likelihood of an area being a potential food desert.