---
title: "Food_Desert_Analysis_2"
author: "Team 1"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
editor_options: 
  markdown: 
    wrap: sentence
---
## Data Loading and Preprocessing
```{r init, include=F}
library(tidyverse)
library(ezids)
library(usmap)
library(ModelMetrics)
library(pROC) 
library(ggplot2)
loadPkg("ISLR")
loadPkg("tree") 
```


```{r setup, include=FALSE}
# Some of common RMD options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
knitr::opts_chunk$set(warning = F, message = F)
# Can globally set option for number display format.
options(scientific=T, digits = 5) 
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r read_data}
# Loading data
data <- read.csv("food_access_research_atlas.csv")

```

# Summary of the dataset:

```{r code1.1, echo=TRUE}
str(data)
head(data)
summary(data)
```


```{r facotorise_data_prepare_for_model}
model_lila_1_10_var = c("State","County","Urban","GroupQuartersFlag","LowIncomeTracts","lahunv1share","PCTGQTRS","MedianFamilyIncome","lawhite1","lablack1","laasian1","lahisp1","lanhopi1","laomultir1","laaian1","lakids10","lakids1","TractKids","laseniors1","laseniors10","TractKids","TractSeniors","TractWhite","TractBlack","TractAsian","TractNHOPI","TractAIAN","TractOMultir","TractHispanic","TractHUNV","TractSNAP", "PovertyRate")
data$State  <- as.factor(data$State)
data$County  <- as.factor(data$County)
data$Urban <- as.factor(data$Urban)
data$GroupQuartersFlag <- as.factor(data$GroupQuartersFlag)
data$LILATracts_1And10 <-  as.factor(data$LILATracts_1And10)
data$LowIncomeTracts <- as.factor(data$LowIncomeTracts)
data$HUNVFlag= as.factor(data$HUNVFlag)
data$LATracts1 = as.factor(data$LATracts1)
data$LATractsVehicle_20 = as.factor(data$LATractsVehicle_20)
data <- ezids::outlierKD2(data, lahunv1share ,qqplt= TRUE, boxplt= TRUE, rm = TRUE)


``` 

```{r selected_var_hist}



plotHistograms <- function(data, variables) {
  # Check if the variables are in the dataframe
  missing_vars <- setdiff(variables, names(data))
  if (length(missing_vars) > 0) {
    stop("The following variables are not in the dataframe: ", paste(missing_vars, collapse = ", "))
  }

  # Create histograms for each variable
  for (var in variables) {
    if (is.numeric(data[[var]])) {
      print(ggplot(data, aes_string(x = var)) + geom_histogram(bins = 30) + 
              ggtitle(paste("Histogram of", var)) +
              xlab(var) + ylab("Frequency"))
    } else {
      warning(paste("Variable", var, "is not numeric. Histograms require numeric variables."))
    }
  }
}

# Usage example
#variables_to_plot <- c("variable1", "variable2", "variable3") # Replace with your variable names
plotHistograms(data, model_lila_1_10_var)

```
```{r}
library(psych)

# Specify the column names you want to include in the pair plot
#selected_columns <- c("colName1", "colName2", "colName3", "colName4", "colName5", "colName6") # Replace with your actual column names

# Use pairs.panels with column names
pairs.panels(data[model_lila_1_10_var], 
             method = "spearman", # correlation method
             hist.col = "#00AFBB",
             density = FALSE,  # show density plots
             ellipses = FALSE # show correlation ellipses
            )

```  

```{r remove_outliers}
variables_to_check <- c("lahunv1share", "MedianFamilyIncome", "PCTGQTRS") 
removeOutliers <- lapply(variables_to_check, function(var){
    # Ensure ezids package is available
  if (!requireNamespace("ezids", quietly = TRUE)) {
      stop("The 'ezids' package is not installed. Please install it using install.packages('ezids').")
  }
  
  
  if (is.numeric(var)) { # Check if the variable is numeric
      # Construct the formula and call outlierKD2 using eval
      
      data <- ezids::outlierKD2(data, x, qqplt = TRUE, boxplt = TRUE, rm = TRUE)
  } else {
      warning(paste("Variable", var, "is not numeric. Skipping outlier removal for this variable."))
  }
    
    
  return(data)
})

# Usage example
#variables_to_check <- c("lahunv1share", "MedianFamilyIncome", "PCTGQTRS") # Replace with your variables#
#data_cleaned <- removeOutliers(data, variables_to_check)


removeOutliers[[2]]

```

```{r logistic_reg_all}
# Create the formula for the glm model
formula_str <- paste("LILATracts_1And10 ~", paste(model_lila_1_10_var, collapse = " + "))

# Fit the GLM model
LILATracts_1And10_logit1 <- glm(as.formula(formula_str), data = data, family = binomial(link = "logit"))


```

```{r logistic_reg_all_analysis_1}
options(max.print = 1e6)
sum_LILATracts_1And10_logit1 = summary(LILATracts_1And10_logit1)
#sum_LILATracts_1And10_logit1
capture.output(sum_LILATracts_1And10_logit1, file = "models/glm_1_logit_lila1_10_summary.txt")
options(max.print = .Options$max.print)
```

```{r save_model}
saveRDS(LILATracts_1And10_logit1, file = "glm_1_logit_lila1_10.rds")

```

```{r check_model}
str(LILATracts_1And10_logit1$fitted.values)
summary(LILATracts_1And10_logit1$fitted.values)
complete_cases <- complete.cases(data[model_lila_1_10_var])
sum(complete_cases)
sapply(data[model_lila_1_10_var], function(x) sum(is.na(x)))
cleaned_data <- na.omit(data, cols = model_lila_1_10_var)
sapply(cleaned_data[model_lila_1_10_var], function(x) sum(is.na(x)))
```

```{r logistic_reg_all_analysis_2}

factors_LILATracts_1And10_logit1 = exp(coef(LILATracts_1And10_logit1))
factors_LILATracts_1And10_logit1
options(max.print = 1e6)
capture.output(factors_LILATracts_1And10_logit1, file = "models/glm_1_logit_lila1_10_factors.txt")

```


All of the coefficients are statistically significant at the $\alpha$ level of 0.05.  

## Accuracy and confusion matrix  
```{r}

cm = table(cleaned_data$LILATracts_1And10,LILATracts_1And10_logit1$fitted.values>.5)
xkabledply( table(cleaned_data$LILATracts_1And10,LILATracts_1And10_logit1$fitted.values>.5), title = "Confusion matrix from Logit Model" )
```

The total accuracy (using the default cutoff of 0.5) is `r round(100*(cm[1,1]+cm[2,2])/sum(cm), digits=1)`% which is pretty decent.

## Receiver-Operator-Characteristic (ROC) curve and Area-Under-Curve (AUC)

ROC and AUC measures the true positive rate (or sensitivity) against the false positive rate (or specificity). The AUC is always between 0.5 and 1.  

```{r }

cleaned_data$prob=predict(LILATracts_1And10_logit1, type = c("response")) # Add new column of predicted probabilities
h <- roc(LILATracts_1And10~prob, data=cleaned_data)
auc(h)
plot(h)
```

We have here the area-under-curve of `r auc(h)`.  This indicates a decent model fit.




# Stepwise forward feature selection

```{r forward_step_LG}
# Load necessary library
library(stats)
model_lila_1_10_var_stepwise = c("Urban","GroupQuartersFlag","lahunv1share","PCTGQTRS","MedianFamilyIncome","lawhite1","lablack1","laasian1","lahisp1","lanhopi1","laomultir1","laaian1","lakids10","lakids1","TractKids","laseniors1","laseniors10","TractKids","TractSeniors","TractWhite","TractBlack","TractAsian","TractNHOPI","TractAIAN","TractOMultir","TractHispanic","TractHUNV","TractSNAP", "PovertyRate")

# Assuming your data is in a dataframe called 'data' and the response variable is 'response'
# Replace 'response' with your actual response variable name

formula_str_2 <- paste("LILATracts_1And10 ~", paste(model_lila_1_10_var_stepwise, collapse = " + "))

# Fit the GLM model
# Initial model with only the intercept (no predictors)
initial_logit2_model <- glm(LILATracts_1And10 ~ 1, data = cleaned_data, family = binomial(link = "logit"))

# Full model with all potential predictors
full_logit2_model <- glm(as.formula(formula_str_2), data = cleaned_data, family = binomial(link = "logit"))

# Perform stepwise forward selection
stepwise_model <- step(initial_logit2_model, scope = list(lower = initial_logit2_model, upper = full_logit2_model), direction = "forward")

# View the summary of the selected model
summary(stepwise_model)

```
```{r save_step_forward_model}
sum_LILATracts_1And10_stepwise_logit2 = summary(stepwise_model)
capture.output(sum_LILATracts_1And10_stepwise_logit2, file = "models/glm_2_stepwise_logit_lila1_10_summary.txt")
```

```{r logistic_reg_all_analysis_stepwise_forward}

factors_LILATracts_1And10_logit2 = exp(coef(stepwise_model))
factors_LILATracts_1And10_logit2
options(max.print = 1e6)
capture.output(factors_LILATracts_1And10_logit2, file = "models/glm_2_stepwise_logit_lila1_10_factors.txt")

```



# Random Forest for poverty Rate


```{r data_prep_pov_rate}
model_pov_rate_columns =  c("LILATracts_1And10","State","County","Urban","GroupQuartersFlag","LowIncomeTracts","lahunv1share","PCTGQTRS","MedianFamilyIncome","lawhite1","lablack1","laasian1","lahisp1","lanhopi1","laomultir1","laaian1","lakids10","lakids1","TractKids","laseniors1","laseniors10","TractKids","TractSeniors","TractWhite","TractBlack","TractAsian","TractNHOPI","TractAIAN","TractOMultir","TractHispanic","TractHUNV","TractSNAP", "")
```

```{r}
# treefitRpart <- rpart(log(Salary) ~ Years + Hits, data=Hitters) # only 7 terminal/leaf  nodes, c.f. 8 from the tree function above
# treefitRpart <- rpart(log(Salary) ~ Years + Hits, data=Hitters, control = list(maxdepth = 8) ) # still only 7 terminal/leaf nodes. cp =0.01 is default (cost complexity)
treefitRpart <- rpart(log(Salary) ~ Years + Hits, data=Hitters, control = list(maxdepth = 8, cp=0.009) ) # 8 terminal/leaf nodes, but slightly different than the nodes from the tree function above. 
#treefit <- tree(log(Salary) ~ Years + Hits, data=Hitters)
summary(treefitRpart)

formula_str <- paste("PovertyRate ~", paste(model_lila_1_10_var, collapse = " + "))

# Fit the GLM model
LILATracts_1And10_logit1 <- glm(as.formula(formula_str), data = data, family = binomial(link = "logit"))
# summary(treefit)
```

```{r}
fancyRpartPlot(treefitRpart)
# For boring plot, use codes below instead
plot(treefitRpart) 
text(treefitRpart,cex=1) # cex control font size
plot(treefit) 
text(treefit,cex=0.75)
```



```{r}
unloadPkg("rattle") 
unloadPkg("rpart.plot")
unloadPkg("ISLR")
unloadPkg("tree") 
```

