---
title: "Keerthana Writeup - Project2"
author: "Keerthana"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: true
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r include=FALSE}
# The package "ezids" (EZ Intro to Data Science) includes a lot of the helper functions we developed for the course. 
# Some of the frequently used functions are loadPkg(), xkabledply(), xkablesummary(), uzscale(), etc.
library(ezids)
 
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```

```{r base_lib}
loadPkg("ggplot2")
```

## Loading dataset

```{r}
df <- data.frame(read.csv("FoodDesert2015.csv")) 
head(df)
#summary(df)
```

## Median Family Income Analysis

From the previous EDA analysis, we learned that Median Family Income is an important feature contributing to the prevalence of a region being a food desert. To further investigate this, we conducted a t-test over the `MedianFamilyIncome` variable.

```{r include=FALSE}
# Set thresholds for defining potential food deserts
poverty_rate_threshold <- 20  # percent
median_income_threshold <- quantile(df$MedianFamilyIncome, 0.25)  # lower quartile
snap_beneficiary_threshold <- 20  # percent

# Creating a food desert indicator
df$PotentialFoodDesert <- with(df, 
  (PovertyRate >= poverty_rate_threshold) &
  (MedianFamilyIncome <= median_income_threshold) &
  (TractSNAP / POP2010 * 100 >= snap_beneficiary_threshold)
)

# Calculating average median family income for potential food desert areas and non-food desert areas
average_income_food_desert <- mean(df[df$PotentialFoodDesert, "MedianFamilyIncome"])
average_income_non_food_desert <- mean(df[!df$PotentialFoodDesert, "MedianFamilyIncome"])

# Performing a t-test to check if the difference in median family incomes is statistically significant
t_test_result <- t.test(
  df[df$PotentialFoodDesert, "MedianFamilyIncome"],
  df[!df$PotentialFoodDesert, "MedianFamilyIncome"]
)

# Print the results
cat("Average Median Family Income in Potential Food Deserts:", average_income_food_desert, "\n")
cat("Average Median Family Income in Non-Food Desert Areas:", average_income_non_food_desert, "\n")
cat("T-Test Results:\n")
print(t_test_result)
```


### T-Test Results

The t-test was conducted to assess the difference in median family incomes between potential food desert areas and non-food desert areas. Here are the results:

Average Median Family Income in Potential Food Deserts: **$22,544.33**
Average Median Family Income in Non-Food Desert Areas: **$67,927.53**

The t-test resulted in a t-statistic of **-149.61** and a p-value of **< 2.2e-16**. The small p-value suggests a significant difference in median family incomes between potential food desert and non-food desert areas.

**Inference:**

Based on the t-test results, we can conclude that there is a statistically significant difference in median family incomes between potential food desert areas and non-food desert areas. Areas classified as potential food deserts tend to have lower median family incomes.


```{r include=FALSE}
logistic_model <- glm(
  PotentialFoodDesert ~ MedianFamilyIncome,
  data = df,
  family = binomial
)

# Summarize the logistic regression results
summary(logistic_model)

exp(coef(logistic_model))
```

### Logistic Regression Results

A logistic regression model was employed to further investigate the relationship between MedianFamilyIncome and the likelihood of an area being a potential food desert. Here are the results:

The odds ratio for MedianFamilyIncome is approximately **-0.915**. This suggests that for every one-unit increase in `MedianFamilyIncome`, the odds of an area being a potential food desert decrease by a factor of approximately **-0.915**, holding other variables constant.

### Effect of a $1,000 Increase in Median Family Income

To understand the effect of a **$1,000** increase in median family income, we can calculate the change in odds:

Change in Odds = Coefficient for **`MedianFamilyIncome` * 1000**

Change in Odds ≈ **-0.915**

So, a **$1,000** increase in median family income is associated with a decrease in the odds of an area being a potential food desert by a factor of approximately **-0.915**, or a **9.15%** reduction in the odds.

This analysis highlights the importance of `MedianFamilyIncome` in determining the likelihood of an area being classified as a potential food desert. Higher median family incomes are associated with a reduced likelihood of an area being a potential food desert.

## Factorizing Categorical variable

```{r encoding}
df$LILATracts_1And10 <- as.factor(df$LILATracts_1And10)
df$GroupQuartersFlag <- as.factor(df$GroupQuartersFlag)
df$Urban = as.factor(df$Urban)
df$LILATracts_halfAnd10 = as.factor(df$LILATracts_halfAnd10)
df$LILATracts_1And20 = as.factor(df$LILATracts_1And20)
df$LILATracts_Vehicle = as.factor(df$LILATracts_Vehicle)
df$HUNVFlag = as.factor(df$HUNVFlag)
df$LowIncomeTracts = as.factor(df$LowIncomeTracts)
df$LA1and10 = as.factor(df$LA1and10)
df$LAhalfand10 = as.factor(df$LAhalfand10)
df$LA1and20 = as.factor(df$LA1and20)
df$LATracts_half = as.factor(df$LATracts_half)
df$LATracts1 = as.factor(df$LATracts1)
df$LATracts10 = as.factor(df$LATracts10)
df$LATracts20 = as.factor(df$LATracts20)
df$LATractsVehicle_20 = as.factor(df$LATractsVehicle_20)
str(df)
```

********** Slide 8 and 9 ****************

## Correlation analysis between GroupQuarters and Food desert:

### Chi sq test analysis:

chi-square (GOF) test between 2 categorical variables (GroupQuartersFlag and LILATracts_1And10):

-   Null Hypothesis (H0): There is no association between the two categorical variables

-   Alternative Hypothesis (H1): There is an association between the two categorical variables

Significance level $\alpha$ = 0.05

```{r chisq, echo=TRUE}
contingency_table <- table(df$GroupQuartersFlag, df$LILATracts_1And10)

chi_squared_test_result <- chisq.test(contingency_table)
chi_squared_test_result

```

Due to the extremely low p value, H0 is rejected. Therefore, there is a **significant association or correlation** between "GroupQuartersFlag" and "LILATracts_1And10."

### Graph analysis:

```{r GQchart1, echo=TRUE}
GroupQuarters_LILA <- df[ (df$GroupQuartersFlag == 1 & df$LILATracts_1And10 == 1), c("GroupQuartersFlag", "LILATracts_1And10")]
NonGroupQuarters_LILA <- df[df$GroupQuartersFlag == 0 & df$LILATracts_1And10 == 1, c("GroupQuartersFlag", "LILATracts_1And10")]

data1 = data.frame()
Percentage_GroupQuarters_LILA = nrow(GroupQuarters_LILA)/(sum(nrow(GroupQuarters_LILA)+ nrow(NonGroupQuarters_LILA)))*100
Percentage_NonGroupQuarters_LILA = nrow(NonGroupQuarters_LILA)/(sum(nrow(GroupQuarters_LILA)+ nrow(NonGroupQuarters_LILA)))*100
data1 <- rbind(data1, Percentage_GroupQuarters_LILA)
data1 <- rbind(data1, Percentage_NonGroupQuarters_LILA)
GroupQuartersFlag = c(1,0)
data1 <- cbind(data1, GroupQuartersFlag)
colnames(data1) <- c('Percentage', 'GroupQuartersFlag')
```

```{r GQchart2, echo=TRUE}

library(ggplot2)
pie_chart <- ggplot(data1, aes(x = "", y = Percentage, fill = factor(GroupQuartersFlag))) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  labs(title = "Percentage of Group quarters in Food desert regions",
       fill = "GroupQuartersFlag") +
  scale_fill_manual(values = c("0" = "grey", "1" = "black"), labels = c("0", "1")) +
  theme_minimal() +
  theme(legend.position = "bottom")
print(pie_chart)

```

The pie chart clearly illustrates that within food desert regions, the percentage of Group Quarters is significantly lower when compared to Non-Group Quarters.

### Impact of Group Quarters on Food desert - Inference:

```{r GQinference1, echo=TRUE}

model1 <- glm(LILATracts_1And10 ~ PovertyRate + GroupQuartersFlag, family = binomial(link = "logit"), data = df)
xkabledply(model1, title = paste("Logistic Regression :", format(formula(model1)) ))
# summary(model1)
```

GroupQuartersFlag = 1 is associated with a decrease of 0.643663 in the log-odds of the response variable (LILATracts_1And10) being 1. The coefficient is significant (p-value = 8e-06).

```{r GQinference2, results='markup', collapse=F}
expcoeff = exp(coef(model1))
#expcoeff
xkabledply(as.table(expcoeff), title = "Exponential of coefficients in Logit Reg" )
```

The effect of being in GroupQuartersFlag = 1, compared to GroupQuartersFlag = 0, is hurting by a factor of `r format(expcoeff[3],digit=4)`, for the log(odds-ratio).  Any factor less than 1 represents a negative effect.

`Analysis` :

Hence, insights drawn from both graphical representation and logistic regression models converge to indicate that `GroupQuartersflag negatively affect the Food deserts.`

********** Slide 10 ****************

## Impact of GroupQuarters and poverty rate on Food desert:

### Graph analysis:

```{r GQPTYchart1, include=FALSE}

data_outliers <- ezids::outlierKD2(df, PovertyRate, rm = TRUE, boxplt = TRUE, qqplt = TRUE)
GroupQuarters_Poverty <- data_outliers[data_outliers$GroupQuartersFlag == 1, c("GroupQuartersFlag", "PovertyRate")]
NonGroupQuarters_Poverty <- data_outliers[data_outliers$GroupQuartersFlag == 0, c("GroupQuartersFlag", "PovertyRate")]

```

```{r GQPTYchart2, echo=TRUE}

data2 = data.frame()
Percentage_GroupQuarters_Poverty = nrow(GroupQuarters_Poverty)/(sum(nrow(GroupQuarters_Poverty)+ nrow(NonGroupQuarters_Poverty)))*100
Percentage_NonGroupQuarters_Poverty = nrow(NonGroupQuarters_Poverty)/(sum(nrow(GroupQuarters_Poverty)+ nrow(NonGroupQuarters_Poverty)))*100
data2 <- rbind(data2, Percentage_GroupQuarters_Poverty)
data2 <- rbind(data2, Percentage_NonGroupQuarters_Poverty)
GroupQuartersFlag = c(1,0)
data2 <- cbind(data2, GroupQuartersFlag)
colnames(data2) <- c('Percentage', 'GroupQuartersFlag')

pie_chart <- ggplot(data2, aes(x = "", y = Percentage, fill = factor(GroupQuartersFlag))) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  labs(title = "Percentage of Poverty in group quarters",
       fill = "GroupQuartersFlag") +
  scale_fill_manual(values = c("0" = "pink", "1" = "black"), labels = c("0", "1")) +
  theme_minimal() +
  theme(legend.position = "bottom")
print(pie_chart)

```

Analysis of the pie chart reveals that Non-Group Quarters tracts exhibit a higher poverty rate compared to Group Quarters tracts.

### Interaction term between group quarters and poverty rate - inference:

```{r GQPTYinference1, echo=TRUE}

model2 <- glm(LILATracts_1And10 ~ PovertyRate * GroupQuartersFlag, family = binomial(link = "logit"), data = df)
xkabledply(model2, title = paste("Logistic Regression :", format(formula(model2)) ))
# summary(model1)
```

* In Non GroupQuarters (GroupQuartersFlag is 0), one unit increase in PovertyRate is associated with a relatively larger increase (0.0570596) in the log-odds of the response variable being 1.

* In GroupQuarters (GroupQuartersFlag is 1), one unit increase in PovertyRate is associated with smaller increase (0.026) in the log-odds of the response variable being 1.


```{r GQPTYinference2, results='markup', collapse=F}
expcoeff = exp(coef(model2))
#expcoeff
xkabledply(as.table(expcoeff), title = "Exponential of coefficients in Logit Reg" )
```

`Analysis` :  

* Through EDA and consideration of interaction terms, it becomes evident that a one-unit increase in the poverty rate within Non-Group Quarters has a more substantial impact on the prevalence of food deserts compared to a one-unit increase in the poverty rate within Group Quarters.

* Consequently, relationship between PovertyRate and the odds of the Food desert is influenced by the presence of GroupQuartersFlag.


********** Slide 11 ****************

## Impact of GroupQuarters and Urban on Food desert:

### Graph analysis:

```{r GQUrbanchart1, echo=TRUE}

GroupQuarters <- df[df$GroupQuartersFlag == 1, c("GroupQuartersFlag", "Urban")]

bar_chart <- ggplot(GroupQuarters, aes(x = GroupQuartersFlag, fill = Urban)) +
  geom_bar(position = "dodge") +
  labs(title = "Comparison of GroupQuartersFlag and Urban Flag",
       x = "Group Quarters Flag",
       y = "Count") +
  scale_fill_manual(values = c("1" = "lightgreen", "0" = "magenta")) +
  theme_minimal()

print(bar_chart)
```

The bar chart highlights a higher prevalence of Group Quarters in Urban areas than in Rural.

### Interaction term between group quarters and Urban - inference:

```{r GQUrbaninference1, echo=TRUE}

model3 <- glm(LILATracts_1And10 ~ GroupQuartersFlag * Urban, family = binomial(link = "logit"), data = df)
xkabledply(model3, title = paste("Logistic Regression :", format(formula(model3)) ))
# summary(model1)
```

* In non-urban areas, the presence of group quarters (GroupQuartersFlag1) is not statistically significant in predicting the log-odds of the response variable being 1.

* In urban areas, the presence of group quarters (GroupQuartersFlag1) is associated with an increase in the log-odds of the response variable being 1. However, the statistical significance of this effect is not strong based on the p-value.


```{r GQUrbaninference2, results='markup', collapse=F}
expcoeff = exp(coef(model3))
#expcoeff
xkabledply(as.table(expcoeff), title = "Exponential of coefficients in Logit Reg" )
```


The log(odds-ratio) for a tract to be food desert is improve by a factor of `r format(expcoeff[4],digit=6)` for GroupQuartersFlag = 1 when Urban1 is 1 compared to when it is 0.

`Analysis` :

The EDA analysis and inference provides the fact that presence of GroupQuarters have stronger positve impact on Food desert in Urban than in Rural.

## Impact of Urban on Food desert:

```{r Urbaninference1, echo=TRUE}

model4 <- glm(LILATracts_1And10 ~ Urban, family = binomial(link = "logit"), data = df)
xkabledply(model4, title = paste("Logistic Regression :", format(formula(model4)) ))
# summary(model1)
```


```{r Urbaninference2, results='markup', collapse=F}
expcoeff = exp(coef(model4))
#expcoeff
xkabledply(as.table(expcoeff), title = "Exponential of coefficients in Logit Reg" )
```

`Analysis`

The effect of a tract being food desert from Urban area, compared to Rural, is associated with a statistically significant increase by a factor of `r format(expcoeff[2],digit=4)`, for the log(odds-ratio).

********** Slide 11 ****************

## Logg odds ratio of Food desert model.

```{r Varimp, echo=TRUE}

variable_values <- c(9.9995e-01, 4.3839e+01, 1.0096e+00, 1.0012e+00, 
                     9.9130e-01, 1.0035e+00, 9.9986e-01, 1.0004e+00, 9.9966e-01, 1.0172e+00,
                     1.0005e+00, 9.5884e-02, 9.9924e-01, 1.0006e+00, 1.0006e+00, 1.9875e+00,
                     1.0005e+00, 1.0010e+00, 9.9960e-01, 1.0001e+00, 9.9904e-01, 1.0005e+00)

variable_names <- c("MedianFamilyIncome", "Urban1", 
                    "laseniors10", "laomultir1", "TractHUNV", "TractSNAP", "lawhite1",
                    "TractWhite", "TractHispanic", "PovertyRate", "lablack1", 
                    "GroupQuartersFlag1", "lakids1", "lahisp1", "TractAIAN", "PCTGQTRS",
                    "TractOMultir", "lakids10", "TractKids", "TractBlack", "laseniors1",
                    "TractSeniors")

# "lahunv1share",    2.0088e+43
merged_df <- data.frame(Variable = variable_names, Importance = variable_values)
merged_df <- merged_df[order(merged_df$Importance), ]

bar_chart <- ggplot(merged_df, aes(x = variable_names, y = variable_values)) +
  geom_bar(stat = "identity", fill = "lightblue", width = 0.5) +
  labs(title = "Log odds ratio of model",
       x = "Variable",
       y = "Log odds") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.2))  

options(
  repr.plot.width = 25,  # Set the desired width in inches
  repr.plot.height = 30  # Set the desired height in inches
)

print(bar_chart)
```

First 4 variables that impact Food desert:

* lahunv1share - 2.0088e+43
* Urban1       - 4.3839e+01
* PCTGQTRS     - 1.9875e+00
* PovertyRate  - 1.0172e+00


********** Slide 18 ****************

## Comparison of CART and Logistic Metrics.

```{r comp, echo=TRUE}

metrics_data <- data.frame(
  Method = c("CART", "CART", "CART", "Logistic", "Logistic", "Logistic"),
  Metric = c("precision", "recall", "accuracy", "precision", "recall", "accuracy"),
  Value = c(0.43029, 0.83996, 0.89117, 0.30786, 0.87071, 0.82149)
)

# Create the line graph using ggplot2
ggplot(metrics_data, aes(x = Metric, y = Value, group = Method, color = Method)) +
  geom_line() +
  geom_point() +
  labs(title = "Comparison of CART and Logistic Metrics",
       x = "Metric",
       y = "Value") +
  theme_minimal()

```


********** Slide 20 ****************

## Impact of Low income tracts on Poverty Rate.

### Graph analysis:

```{r GQUrbanchart1, include=FALSE}

LowIncomeTracts_Poverty <- df[df$LowIncomeTracts == 1, c("LowIncomeTracts", "PovertyRate")]
NonLowIncomeTracts_Poverty <- df[df$LowIncomeTracts == 0, c("LowIncomeTracts", "PovertyRate")]

```

```{r GQUrbanchart2, echo=TRUE}

data2 = data.frame()
Percentage_LowIncomeTracts_Poverty = sum(LowIncomeTracts_Poverty$PovertyRate)/(sum(nrow(LowIncomeTracts_Poverty)+ nrow(NonLowIncomeTracts_Poverty)))*100
Percentage_NonLowIncomeTracts_Poverty = sum(NonLowIncomeTracts_Poverty$PovertyRate)/(sum(nrow(LowIncomeTracts_Poverty)+ nrow(NonLowIncomeTracts_Poverty)))*100
data2 <- rbind(data2, Percentage_LowIncomeTracts_Poverty)
data2 <- rbind(data2, Percentage_NonLowIncomeTracts_Poverty)
LowIncomeTracts = c(1,0)
data2 <- cbind(data2, LowIncomeTracts)
colnames(data2) <- c('Percentage', 'LowIncomeTracts')

library(ggplot2)
ggplot(data2, aes(x = factor(LowIncomeTracts), y = Percentage, fill = factor(LowIncomeTracts))) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c("0" = "lightgreen", "1" = "lightblue")) +
  labs(title = "Percentage of Poverty in LowIncomeTracts",
       fill = "LowIncomeTracts") +
  theme_minimal() +
  theme(legend.position = "bottom")

```

```{r PTYLITinference1, echo=TRUE}

model5 <- lm(PovertyRate ~ LowIncomeTracts, data = df)
xkabledply(model5, title = paste("Logistic Regression :", format(formula(model5)) ))
# summary(model1)
```

In this case, it suggests that when "LowIncomeTracts" is 1 (compared to 0), the estimated poverty rate increases by approximately 18.72349 units.

**The model indicates that the presence of low-income tracts (LowIncomeTracts1) is a statistically significant predictor of the poverty rate.**


********** Slide 21 ****************

## Impact of Food desert on Poverty Rate.

```{r PTYLITinference1, echo=TRUE}

model6 <- lm(PovertyRate ~ LILATracts_1And10, data = df)
xkabledply(model6, title = paste("Logistic Regression :", format(formula(model6)) ))
# summary(model1)
```

It suggests that when "LILATracts_1And10" changes from 0 to 1, the estimated poverty rate increases by approximately 11.35890 units.

**The coefficient for "LILATracts_1And101" suggests a positive association between the presence of such tracts and the poverty rate.**


********** Slide 26 to Slide 30 ****************

## Linear regression assumptions:

### Normality of Median family income

```{r norm1, echo=TRUE}

hist(df$MedianFamilyIncome, main = "Histogram of Median Family Income", xlab = "Median Family Income")

# Check normality with a Q-Q plot
qqnorm(df$MedianFamilyIncome, main = "Q-Q Plot of Median Family Income")
qqline(df$MedianFamilyIncome, col = 2)
```

**Removing outliers**

```{r norm2 ,echo=TRUE}
MedianFamilyIncome_outliers <- ezids::outlierKD2(df, MedianFamilyIncome, rm = TRUE, boxplt = TRUE, qqplt = TRUE)
```

### Normality of Poverty rate:

```{r norm3, echo=TRUE}
hist(df$PovertyRate, main = "Histogram of PovertyRate", xlab = "PovertyRate")

# Check normality with a Q-Q plot
qqnorm(df$PovertyRate, main = "Q-Q Plot of PovertyRate")
qqline(df$PovertyRate, col = 2)
```
**Removing outliers**

```{r norm2 ,echo=TRUE}
PovertyRate_outliers <- ezids::outlierKD2(df, PovertyRate, rm = TRUE, boxplt = TRUE, qqplt = TRUE)
```


********** Slide 24 ****************


## Variable importance of Linear regression model.

```{r varimp2, echo=TRUE}

variable_values <- c(100, 58.0798, 38.5578, 29.0375, 12.6898, 11.8240, 9.7042, 0, 0, 0)

variable_names <- c("MedianFamilyIncome","LowIncomeTracts","TractSNAP","TractHUNV","TractBlack","TractSeniors","TractWhite", "lahisp","lawhite1","lakids10")
  
merged_df <- data.frame(Variable = variable_names, Importance = variable_values)
merged_df <- merged_df[order(merged_df$Importance), ]

bar_chart <- ggplot(merged_df, aes(x = variable_names, y = variable_values)) +
  geom_bar(stat = "identity", fill = "lightblue", width = 0.5) +
  labs(title = "Variable importance",
       x = "Variable",
       y = "importance") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.2))  

options(
  repr.plot.width = 25,  # Set the desired width in inches
  repr.plot.height = 30  # Set the desired height in inches
)

print(bar_chart)
```

Top 4 variables impacting Poverty rate:

* MedianFamilyIncome   100.0000
* LowIncomeTracts       58.0798
* TractSNAP             38.5578
* TractHUNV             29.0375

    
********** Slide 25 ****************

## Comparison of CART and Linear Regression Metrics.

```{r comp2, echo=TRUE}
metrics_data <- data.frame(
  Method = c("Linear", "Linear", "CART", "CART"),
  Metric = c("R2", "RMSE", "R2", "RMSE"),
  Value = c(0.63682, 7.4614, 0.68551, 6.9420)
)

# Create the line graph using ggplot2
ggplot(metrics_data, aes(x = Metric, y = Value, group = Method, color = Method)) +
  geom_line() +
  geom_point() +
  labs(title = "Comparison of CART and Linear Metrics",
       x = "Metric",
       y = "Value") +
  theme_minimal()

```


