---
title: "Food Deserts Analysis"
author: "Team 1"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes

---
## Data Loading and Preprocessing
```{r init, include=F}
library(tidyverse)
library(ezids)
library(usmap)
```


```{r setup, include=FALSE}
# Some of common RMD options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
knitr::opts_chunk$set(warning = F, message = F)
# Can globally set option for number display format.
options(scientific=T, digits = 5) 
```

```{r read_data}
# Loading data
data <- read.csv("food_access_research_atlas.csv")

```

## Questions 1a: Geographic Distribution of State/County wise Prevelance of Food Deserts

```{r}
plot_map <- function(dataset, colx, region, subx) {
  grpddata <- dataset %>%
    filter(.data[[colx]] == 1) %>% # Filtering the food desert flags
    group_by(State, County) %>% # Grouping by State and County
    reframe(full = State, count = sum(.data[[colx]]))  # Reframing to dataframe for Plotting
  grpddata$county <-
    paste(grpddata$County, "County") # Adding " County" text to support prerequisite of the plot_usmap() counties
  grpddata <-
    distinct(grpddata, .keep_all = TRUE) %>% select(full, county, count) # Ignoring the duplicates and selecting particular columns to supply the plot_usmap()
  counties_df <-
    us_map("counties") %>% select(fips, full, county) %>% distinct() # Getting the FIPS data of the counties. Mandatory for plotting in maps. Available from the us_map() dataframe built-in with the plot_usmap() library
  merged_df <-
    grpddata %>% right_join(counties_df, by = c("county", "full")) # Right joining the dataframes to get relevant FIPS code
  
  if (region == "State") {
    merged_df <-
      merged_df %>% select(-fips) %>% rename(state = full) # Changing the structure for State wise plotting
  }
  state_map <-
    plot_usmap("states", color = "#ff000030", size = 0.01)
  
  counties_map <- plot_usmap(
    data = merged_df,
    values = "count",
    color = "black",
    size = 0.1
  )
  
  # Merging both the plots using ggplot and rendering the combined map using geom_polygon()
  ggplot() +
    geom_polygon(
      data = counties_map[[1]],
      aes(
        x = x,
        y = y,
        group = group,
        fill = counties_map[[1]]$count,
      ),
      color = "black",
      size = 0.1
    ) +
    geom_polygon(
      data = state_map[[1]],
      aes(x = x,
          y = y,
          group = group),
      color = "#ff000030",
      fill = alpha(0.01)
    ) +
    coord_equal() +
    theme(
      panel.grid = element_blank(),
      axis.title = element_blank(),
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      strip.text = element_blank(),
      plot.subtitle = element_text(size = rel(0.8)),
      panel.background = element_rect(fill = "white"),
      
    ) +
    scale_fill_gradient(low = 'white', high = 'grey20') +
    labs(
      title = paste(region, "- Wise Geographic Distribution of Food Desert in US"),
      subtitle = subx,
      fill = "Prevelance of FD"
    )
}


plot_map(data, colx = "LILATracts_halfAnd10", region="County", subx="LI and LA half and 10 miles for Urban and Rural")
plot_map(data, colx = "LILATracts_1And10", region="County", subx="LI and LA 1 and 10 miles for Urban and Rural")
plot_map(data, colx = "LILATracts_1And20", region="County", subx="LI and LA 1 and 20 miles for Urban and Rural")
plot_map(data, colx = "LILATracts_halfAnd10", region="State", subx="LI and LA half and 10 miles for Urban and Rural")
plot_map(data, colx = "LILATracts_1And10", region="State", subx="LI and LA 1 and 10 miles for Urban and Rural")
plot_map(data, colx = "LILATracts_1And20", region="State", subx="LI and LA 1 and 20 miles for Urban and Rural")
```

## Question 1b: Unique states and counties are represented in the dataset

```{r}
unique_states <- data %>% group_by(State) %>% distinct(State) # Grouping by State and using distinct() to get unique states
unique_counties <- data %>% group_by(State, County) %>% distinct(County) 
# Grouping by both State and County and then 
# using distinct() to get unique states. Because directly Grouping the counties will result in lesser results as different 
# counties in different States can exists with the same name
```
## Question 1c: Total census tracts included in the dataset
```{r}
total_census_tracts <- nrow(data)
```

```{r}
unique_states <- data %>% group_by(State) %>% distinct(State) # Grouping by State and using distinct() to get unique states
unique_counties <- data %>% group_by(State, County) %>% distinct(County) 
# Grouping by both State and County and then 
# using distinct() to get unique states. Because directly Grouping the counties will result in lesser results as different 
# counties in different States can exists with the same name
```

## Question 2: Urban vs Rural Classification impacts
```{r}
total_count <- nrow(data)
urban_count <- nrow(data %>% filter(Urban == 1))
rural_count <- nrow(data %>% filter(Urban == 0))
urban_food_desert_count <- nrow(data %>% filter(Urban == 1, LILATracts_1And10 == 1))
rural_food_desert_count <- nrow(data %>% filter(Urban == 0, LILATracts_1And10 == 1))

urban_percentage <- urban_count / total_count * 100
rural_percentage <- rural_count / total_count * 100
urban_food_desert_percentage <- urban_food_desert_count / urban_count * 100
rural_food_desert_percentage <- rural_food_desert_count / rural_count * 100
overall_percentage <- (urban_food_desert_count + rural_food_desert_count) / total_count * 100

cat("There are", urban_count, "Urban Counties\n")
cat("There are", rural_count, "Rural Counties\n\n")

cat("There are", urban_food_desert_count, "Urban Food Deserts\n")
cat("There are", rural_food_desert_count, "Rural Food Deserts\n\n")

cat("The Urban tracts constitute", round(urban_percentage, 2), "% of the total Census tracts\n")
cat("The Rural tracts constitute", round(rural_percentage, 2), "% of the total Census tracts\n\n")

cat("Based on the dataset, it is found that", round(urban_food_desert_percentage, 2), "% of the Urban tracts are food deserts\n")
cat("And", round(rural_food_desert_percentage, 2), "% of the Rural tracts are food deserts\n\n")

cat("Overall,", round(overall_percentage, 2), "% are food deserts based on the analysis made over the dataset\n")
cat("From the analysis, it is evident that Food Deserts are more common in the Urban Areas, due to many factors like Population, Poverty Rates, Low Access and Low Income tracts")
```

## Question 4: Relationship between Poverty Rate and Food Deserts

### Analyzing Poverty Rate in Different Tracts

```{r}

# Assuming `data` is the dataframe you have
data %>% 
  group_by(LILATracts_1And10) %>% 
  summarise(mean_poverty_rate = mean(PovertyRate))

data %>% 
  group_by(LILATracts_1And20) %>% 
  summarise(mean_poverty_rate = mean(PovertyRate))

data %>% 
  group_by(LowIncomeTracts) %>% 
  summarise(mean_poverty_rate = mean(PovertyRate))

## Poverty rate is high in non low access areas
data %>% 
  group_by(LA1and10) %>% 
  summarise(mean_poverty_rate = mean(PovertyRate))

data %>% 
  group_by(LA1and20) %>% 
  summarise(mean_poverty_rate = mean(PovertyRate))

```

The mean poverty rate is 26.5 for food deserts and for non-food deserts it is 15.1. So the Poverty rate in food deserts is 75% higher than non-food deserts. Poverty rate in Low access area is 14.2 whereas in non-low access area is 18.0.

### Count of Low Income tracts in Low access tracts
```{r low_income_low_access}

# Filtering data for rows where LA1and10 is 1
filtered_data <- data[data$LA1and10 == 1, ]

# Calculating value counts for LowIncomeTracts
value_counts <- table(filtered_data$LowIncomeTracts)

# Converting to data frame for ggplot
value_counts_df <- as.data.frame(value_counts)
names(value_counts_df) <- c("LowIncomeTracts", "Counts")

# Creating custom labels
value_counts_df$Labels <- ifelse(value_counts_df$LowIncomeTracts == 0, "Not Low Income", 
                                 ifelse(value_counts_df$LowIncomeTracts == 1, "Low Income", value_counts_df$LowIncomeTracts))

# Creating the bar plot
p <- ggplot(value_counts_df, aes(x = reorder(Labels, -Counts), y = Counts)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  labs(title = "Counts of LowIncomeTracts in low access areas", x = "LowIncomeTracts", y = "Counts") +
  theme(axis.text.x = element_text(angle = 0, hjust = 1))

# Showing the plot
print(p)

```

* 66% of the non-low income tracts are low access tracts
* This means that Low access people are rich which may indicate that rich people wants to stay outside.


## Relationship between states and poverty rate

### Statewise average poverty rate

```{r statewise_poverty_rate }
# 
# states_pov <- data %>%
#   group_by(State) %>%
#   summarise(avg_poverty_rate = mean(PovertyRate))
# #print(states_pov)


states_pov <- data %>%
  group_by(State) %>%
  summarise(avg_poverty_rate = mean(PovertyRate, na.rm = TRUE)) %>%
  arrange(desc(avg_poverty_rate))
print(states_pov)
print(sum(is.na(data$PovertyRate)))
```
* Mississippi has the highest povery rate of 24.64 and Florida has the lowest poverty rate of 24.64.
* There are no missing values in the PovertyRate.

### Normalize poverty rate

```{r statewise_poverty_rate_normalise }
states_pov$norm_poverty_rate <- scale(states_pov$avg_poverty_rate)
print(states_pov)
```

The poverty rate is normalised to find the correlation between poverty rate of states and food deserts

### Identify states with high poverty
```{r high_pov_states}
threshold <- quantile(states_pov$norm_poverty_rate, 0.85)
states_high_pov <- states_pov %>% 
  filter(norm_poverty_rate > threshold)
print(states_high_pov)

```
To address the question whether states with high poverty rate also have high number of food deserts, we are filtering only the states above 85th percentile of poverty rate.

### Correlation between states and food deserts

```{r corr_states_food_deserts}
# Calculate statewise food desert count
states_food_deserts <- data %>% 
  filter(LILATracts_1And10 == 1) %>% 
  count(State)
#print(states_food_deserts)

# Normalize food desert counts
states_food_deserts$norm_count <- scale(states_food_deserts$n)

merged_data <- inner_join(states_high_pov, states_food_deserts, by = "State")
merged_data_all <- inner_join(states_pov, states_food_deserts, by = "State")
# Get correlation
# There is no correlation between poverty rates and food desert status
# The states with high poverty don't have high number of food deserts
#print(cor(merged_data$norm_poverty_rate, merged_data$norm_count, method = "pearson"))
#### Checking correlation of poverty rate and food desert status of all the states 
#### There is very less correlation between povery rate and food deserts in all states
#print(cor(states_pov$norm_poverty_rate, states_food_deserts$norm_count, method = "pearson"))

#print(merged_data)
cor.test(merged_data$norm_poverty_rate, merged_data$norm_count, method="pearson")
#cor.test(states_pov$norm_poverty_rate, states_food_deserts$norm_count, method="pearson")
cor.test(merged_data_all$norm_poverty_rate, merged_data_all$norm_count, method="pearson")

```
### Pearson Correlation:

#### High Poverty States and food desert count

**Null hypothesis**: There is no correlation between High poverty rate states with count of their food desert
**Alternate hypothesis** : There is Correlation between High poverty rate states with count of their food desert.
At level of significance alpha set at = 0.05, as The p-value 0.6  is higher than the 0.05, we fail to reject the null hypothesis.
The pearson correlation results tell us "There is no correlation between High poverty rate states with count of their food desert"

#### Poverty rate of All States and food desert count

**Null hypothesis**: There is no correlation between poverty rate of all states with count of their food desert
**Alternate hypothesis** : There is Correlation between poverty rate of all states with count of their food desert.
At level of significance alpha set at = 0.05, as The p-value 0.003  is lesser than the 0.05, we reject the null hypothesis.
The pearson correlation results tell us there is statistically significant correlation between poverty rate of all states with count of their food desert.
 

## Relationship between counties and poverty rate
We calculate the correlation between counties and poverty rate just like we did for the states.

### Calculate countywise average poverty rate

```{r countywise_poverty_rate }

counties_pov <- data %>% 
  group_by(County) %>% 
  summarise(avg_poverty_rate = mean(PovertyRate,na.rm = TRUE)) %>%
  arrange(desc(avg_poverty_rate))
print(head(counties_pov))
print(tail(counties_pov))
```

* Radford has the highest average povery rate of 24.64 and Bedford City has the lowest average poverty rate of 24.64.

### Normalize county poverty rate

```{r countywise_poverty_rate_normalise }
counties_pov$norm_poverty_rate <- scale(counties_pov$avg_poverty_rate)
print(head(counties_pov))
```
### Identify county with high poverty
```{r high_pov_counties}
county_threshold <- quantile(counties_pov$norm_poverty_rate, 0.85)
counties_high_pov <- counties_pov %>% 
  filter(norm_poverty_rate > county_threshold)
print(head(counties_high_pov))

```

### Correlation between counties and food deserts

```{r corr_counties_food_deserts}
# Calculate statewise food desert count
counties_food_deserts <- data %>% 
  filter(LILATracts_1And10 == 1) %>% 
  count(County)
#print(counties_food_deserts)

# Normalize food desert counts
counties_food_deserts$norm_count <- scale(counties_food_deserts$n)

#length(counties_food_deserts$norm_poverty_rate)
#length(counties_food_deserts$norm_count)

counties_merged_data <- inner_join(counties_high_pov, counties_food_deserts, by = "County")

# Get correlation
# There is no correlation between poverty rates and food desert status
# The states with high poverty don't have high number of food deserts
#print(cor(counties_merged_data$norm_poverty_rate, counties_merged_data$norm_count, method = "pearson"))
#### Checking correlation of poverty rate and food desert status of all the states 
#### There is very less correlation between povery rate and food deserts in all states
counties_all_merged_data <- inner_join(counties_pov, counties_food_deserts, by = "County")
#print(cor(counties_all_merged_data$norm_poverty_rate, counties_all_merged_data$norm_count, method = "pearson"))

cor.test(counties_merged_data$norm_poverty_rate, counties_merged_data$norm_count, method="pearson")
cor.test(counties_all_merged_data$norm_poverty_rate, counties_all_merged_data$norm_count, method="pearson")


# print(sum(is.na(counties_all_merged_data$norm_poverty_rate)))
# sum(is.infinite(counties_all_merged_data$norm_poverty_rate))
# print(sum(is.na(counties_all_merged_data$norm_count)))
# sum(is.infinite(counties_all_merged_data$norm_count))
#print(counties_all_merged_data)
```

### Pearson Correlation:

#### High Poverty counties and food desert count

**Null hypothesis**: There is no correlation between High poverty rate counties with count of their food desert
**Alternate hypothesis** : There is Correlation between High poverty rate counties with count of their food desert.
At level of significance alpha set at = 0.05, as The p-value 1  is higher than the 0.05, we fail to reject the null hypothesis. The correlation coefficient is cor 5.98e-05 .
The pearson correlation results tell us "There is no correlation between High poverty rate counties with count of their food desert"

#### Poverty rate of All counties and food desert count

**Null hypothesis**: There is no correlation between poverty rate of all counties with count of their food desert.
**Alternate hypothesis** : There is Correlation between poverty rate of all counties with count of their food desert.
At level of significance alpha set at = 0.05, as The p-value 0.003  is lesser than the 0.05, we reject the null hypothesis.The correlation coefficient is cor 0.0147. 
The pearson correlation results tell us there is no statistically significant correlation between poverty rate of all counties with count of their food desert.

# Question 5: The impact of vehicle availability on food desert designation

## Visualizing Vehicle Access

#### Filtering and plotting vehicle access

```{r vis_vehicle_access}
data %>%
  filter(LILATracts_1And10 == 1) %>%
  ggplot(aes(x = lahunv1share)) +
  geom_histogram(bins = 50,fill = "skyblue", color="darkred") +
  geom_density(alpha = 0.5,fill= "red", color="darkred") +
  labs(title = "Percentage of total population who does not have vehicles in food desert")

data %>%
  filter(LILATracts_1And10 == 0) %>%
  ggplot(aes(x = lahunv1share)) +
  geom_histogram(bins = 50,fill = "skyblue", color="darkred") +
  geom_density(alpha = 0.5,fill= "red", color="darkred") +
  labs(title = "Percentage of total population who does not have vehicles in non-food desert")

data %>%
  filter(LILATracts_1And10 == 1 & Urban == 1) %>%
  ggplot(aes(x = lahunv1share)) +
  geom_histogram(bins = 50,fill = "skyblue", color="darkred") +
  geom_density(alpha = 0.5,fill = "red", color="darkred") +
  labs(title = "Percentage of urban population who does not have vehicles in food desert")

data %>%
  filter(LILATracts_1And10 == 1 & Urban == 0) %>%
  ggplot(aes(x = lahunv1share)) +
  geom_histogram(bins = 50,fill = "skyblue", color="darkred") +
  geom_density(alpha = 0.5,fill = "red", color="darkred") +
  labs(title = "Percentage of rural population who does not have vehicles in food desert")

data %>%
  filter(LILATracts_1And10 == 0 & Urban == 1) %>%
  ggplot(aes(x = lahunv1share)) +
  geom_histogram(bins = 50,fill = "skyblue", color="darkred") +
  geom_density(alpha = 0.5,fill = "red", color="darkred") +
  labs(title = "Percentage of urban population who does not have vehicles in non-food desert")

data %>%
  filter(LILATracts_1And10 == 0 & Urban == 0) %>%
  ggplot(aes(x = lahunv1share)) +
  geom_histogram(bins = 50,fill = "skyblue", color="darkred") +
  geom_density(alpha = 0.5,fill = "red", color="darkred") +
  labs(title = "Percentage of rural population who does not have vehicles in non-food desert")

# fd_urban <- data %>%
#   filter(LILATracts_1And10 == 1 & Urban == 1) %>%
#   summarise(per_without_vehicle_urban = mean(lahunv1share,na.rm = TRUE))
# print(fd_urban)
#
# fd_rural <- data %>%
#   filter(LILATracts_1And10 == 1 & Urban == 0) %>%
#    summarise(per_without_vehicle_rural = mean(lahunv10share,na.rm = TRUE))
# print(fd_rural)

# fd_vehicle <- data %>%
#   filter(LILATracts_1And10 == 1) %>%
#    summarise(per_with_vehicle = mean(lahunv10share,na.rm = TRUE))
# print(fd_vehicle)

result <- data %>%
  filter(LILATracts_1And10 == 1) %>%
  group_by(Urban) %>%
  summarise(
    lahunv1share_mean = mean(lahunv1share, na.rm = TRUE),
    lahunv10share_mean = mean(lahunv10share, na.rm = TRUE)
  )

print(result)
fd_urban = subset(result, Urban == 1)["lahunv1share_mean"] * 100
fd_rural = subset(result, Urban == 0)["lahunv10share_mean"] * 100
fd_vehicle = 100 - fd_urban + fd_urban

```

```{r vehicle_access_pie}
#print(result[[2, "lahunv1share_mean"]])
#print(fd_urban)
data_pie <- data.frame(
  Category = c("Urban without vehicles", "Rural without vehicles", "Population with Vehicles"),
  Percentage = c(result[[2, "lahunv1share_mean"]] * 100 ,result[[1, "lahunv10share_mean"]] * 100, 100 - (result[[1, "lahunv1share_mean"]]  + result[[2, "lahunv10share_mean"]])* 100)
)
#print(data_pie)
# Generate the pie chart
ggplot(data_pie, aes(x = "", y = Percentage, fill = Category)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start=0) +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.border = element_blank(),
    panel.grid=element_blank(),
    axis.ticks = element_blank(),
    plot.title=element_text(size=14, face="bold")
  ) +
  labs(fill = "") +
  scale_fill_manual(values = c("skyblue", "darkred", "pink"))
```

Around 5% urban population in food desert doesn't have vehicles. Around 4% rural population in food desert doesn't have vehicles. 91% population in food desert have vehicles.

### Share of population without vehicles - boxplots

``` {r vehicle_box_plots}
data$LILATracts_1And10 <- as.factor(data$LILATracts_1And10)


levels(data$LILATracts_1And10) <- c("Non-Food desert", "Food desert")

ggplot(data = data,
       aes(x = LILATracts_1And10, y =lahunv1share)) +
      geom_boxplot(fill='white', color="darkred") +
      labs(title="Percentage of population doesn't have vehicles in food desert and non-food deserts",
      x ="Types of tracts ", y = "Population share without vehicles") +
      theme(plot.title = element_text(hjust = 0.5))

```

Seems like there is a lot of outliers in the data to make any conclusions. So removed outliers using KD2 method.

```{r outlier_removal_vehicle_access_food_deserts}
app_wo_gre_outlier <- ezids::outlierKD2(data, lahunv1share ,qqplt= TRUE, boxplt= TRUE, rm = TRUE)
#str(app_wo_gre_outlier)
#gre_na = sum(is.na(app_wo_gre_outlier$gre))
```

The QQ plot looks like a straight line after removing outliers. This indicates that it is normally distributed.

``` {r vehicle_box_plots_2}
app_wo_gre_outlier$LILATracts_1And10 <- as.factor(app_wo_gre_outlier$LILATracts_1And10)


levels(app_wo_gre_outlier$LILATracts_1And10) <- c("Non-Food desert", "Food desert")

ggplot(data = app_wo_gre_outlier,
       aes(x = LILATracts_1And10, y =lahunv1share)) +
      geom_boxplot(fill='white', color="darkred") +
      labs(title="Percentage of total population who does not have vehicles in food desert and non-food deserts",
      x ="Types of tracts ", y = "Population share without vehicles") +
      theme(plot.title = element_text(hjust = 0.5))

```

* A higher percentage of the population in food deserts lack vehicles compared to those in non-food desert.
* This could be concerning, as people in food deserts without vehicles may have even more difficulty accessing fresh and healthy food.

#### Observations from the box plot

Central Tendency: The median line inside the box for the "food desert" category appears to be higher than the "non-food desert" category. This suggests that, on average, a higher percentage of the population in food deserts lack vehicles compared to those in non-food deserts.

Spread & Variability: The interquartile range is higher for food desert.

Outliers: There are some outliers in both categories, but it's particularly noticeable in the "non-food desert" category. This indicates that while most non-food desert tracts have a relatively lower percentage of people without vehicles, there are a few tracts where this isn't the case.

Skewness: The median line appears to be roughly in the middle of the box for food desert suggesting that the distribution of the share of the population without vehicles in food desert is roughly symmetric, but non-food desert is skewed towards the right.

Overall Comparison: The entire box (representing the middle 50% of the data) for the "food desert" category is higher on the y-axis compared to the "non-food desert" category. This indicates that a larger share of the population in food deserts typically lacks vehicles compared to those in non-food deserts. This could be concerning, as people in food deserts without vehicles may have even more difficulty accessing fresh and healthy food.

### Vehicle access in rural and urban tracts

```{r veh_access_urban_rural}
data$Urban <- as.factor(data$Urban)
levels(data$Urban) <- c("Rural", "Urban")
ggplot(data = data,
       aes(x = interaction(LILATracts_1And10, Urban), y = lahunv1share)) +
  geom_boxplot(fill='white', color="darkred") +
  labs(title="Percentage of total population who does not have vehicles",
       x ="Types of tracts", y = "Population share without vehicles") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1))
```

Seems like there is a lot of outliers in the data to make any conclusions. So let's use outlier removed data.


```{r veh_access_urban_rural_without_outliers}
app_wo_gre_outlier$Urban <- as.factor(app_wo_gre_outlier$Urban)
levels(app_wo_gre_outlier$Urban) <- c("Rural", "Urban")

app_wo_gre_outlier$LILATracts_1And10 <- as.factor(app_wo_gre_outlier$LILATracts_1And10)
levels(app_wo_gre_outlier$LILATracts_1And10) <- c("Non-food desert", "Food desert")
ggplot(data = app_wo_gre_outlier,
       aes(x = interaction(LILATracts_1And10, Urban), y = lahunv1share)) +
  geom_boxplot(fill='white', color="darkred") +
  labs(title="Percentage of total population who does not have vehicles - outliers removed",
       x ="Types of tracts", y = "Population share without vehicles") +
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1))
```

#### Non-Food Desert in Urban (3rd category):
* Central Tendency: The median suggests that a relatively low percentage of the population in non-food desert urban areas lack vehicle access, even lower than the non-food desert rural areas.

#### Urban Areas:
The non-food desert urban areas have the smallest median percentage of people without vehicle access. This is likely because urban areas often have alternative modes of transportation, such as public transit, walking, or biking, making vehicle ownership less necessary. Moreover, urban areas often have amenities and services within walking distance, further reducing the need for a vehicle.

#### Impact of Food Deserts:
In both rural and urban contexts, food deserts show a higher median percentage of people without vehicle access compared to non-food deserts. This is a significant concern because it suggests that people in food deserts, who already might have limited access to healthy food options, might also face transportation challenges.

#### Rural vs. Urban:
While non-food desert urban areas have the lowest median percentage of people without vehicle access, food desert rural areas have the highest. This highlights the impact of food desert status on transportation barriers, irrespective of the urban or rural context.



# Question 6: Income-related Variables and Food Deserts
## Visualizing Median Family Income

```{r medincom_food_deserts}


data %>%
  filter(LILATracts_1And10 == "Food desert") %>%
  ggplot(aes(x = MedianFamilyIncome)) +
  geom_histogram(bins = 50, fill='skyblue', color="darkred") +  # Changed colors here
  geom_density(alpha = 0.8, fill="red", color="darkred") +      # And here
  labs(title = "Median Family Income in Food Deserts")

data %>%
  filter(LILATracts_1And10 == "Non-Food desert") %>%
  ggplot(aes(x = MedianFamilyIncome)) +
  geom_histogram(bins = 50, fill='skyblue', color="darkred") +  # Changed colors here
  geom_density(alpha = 0.5, fill = "red", color="darkred") +      # And here
  labs(title = "Median Family Income in Non-food Deserts")

```


1.	**Median Family Income in Food Deserts**: This histogram appears to be skewed to the right, indicating that a significant portion of families in food deserts have lower incomes. There seems to be a peak around the lower income values, which tapers off as income increases.
2.	**Median Family Income in Non-Food Deserts**: This histogram looks more symmetrical and possibly bell-shaped, suggesting a more normal distribution. The peak is closer to the middle income values, and the tails decrease symmetrically on both sides.
3.	From this visual inspection, it seems that food deserts tend to have a higher proportion of families with lower incomes, while non-food deserts have a more even distribution of incomes, with a peak around the middle values.


### Median income in food desert - boxplots

``` {r medincome_box_plots}
data$LILATracts_1And10 <- as.factor(data$LILATracts_1And10)


levels(data$LILATracts_1And10) <- c("Non-Food desert", "Food desert")

ggplot(data = data,
       aes(x = LILATracts_1And10, y = MedianFamilyIncome)) +
      geom_boxplot(fill='white', color="darkred") +
      labs(title="Median Family Income between food desert and non-food deserts",
      x ="Types of tracts ", y = "MedianFamilyIncome") +
      theme(plot.title = element_text(hjust = 0.5))

```


## Confidence Interval tests

```{r food_desert_confidence_interval}
fd_data <- subset(data, data$LILATracts_1And10 == "Food desert")

non_fd_data <- subset(data, data$LILATracts_1And10 == "Non-Food desert")
#print(fd_data)
ttest_fd_medincome_95 = t.test(x=fd_data$MedianFamilyIncome, conf.level=0.95 )
ttest_non_fd_medincome_95 = t.test(x=non_fd_data$MedianFamilyIncome, conf.level=0.95 )

print(ttest_fd_medincome_95)
print(ttest_non_fd_medincome_95)
```

#### Food deserts

The T-intervals for the median income

* **At 0.95 level**

   [`r ttest_fd_medincome_95$conf.int[1] `, `r ttest_fd_medincome_95$conf.int[2] `]

#### Non-Food deserts

The T-intervals for the median income in non-food desert

* **At 0.95 level**

   [`r ttest_non_fd_medincome_95$conf.int[1] `, `r ttest_non_fd_medincome_95$conf.int[2] `]

The difference in the intervals indicates the median income between food desert and non-food desert is different.

## ANOVA ON median income
Null hypothesis : There is no difference in the median income between food desert and non-food desert
Alternate hypothesis : There is difference in the median income between food desert and non-food desert


```{r anova_med_incom_food_desert}
med_income_aov = aov(MedianFamilyIncome ~ LILATracts_1And10, data = data)
# plant_aov  # this does not give the easy-to-read result of the aov analysis
# summary(plant_aov)
plantaovsummary = summary(med_income_aov)
plantaovsummary
```

At level of significance alpha set at = 0.05, as The p-value 2e-16  is less than the 0.05, we reject the null hypothesis "There is no difference in the median income between food desert and non-food desert"
Given the large F value and the extremely low p-value, we can be confident in this result.

The anova results tell us that there is statistically significant difference in median income among food desert and non -deserts.


### Post-hoc Tukey HSD

```{r tukeyHSD}
tukeyAoV <- TukeyHSD(med_income_aov)
tukeyAoV
```

The post-hoc test confirms that families living in food deserts have a significantly lower median income compared to those not living in food deserts.

#### Median Family Income in urban and rural areas

``` {r medincome_urban_box_plots}
data$LILATracts_1And10 <- as.factor(data$LILATracts_1And10)
levels(data$LILATracts_1And10) <- c("Non-Food desert", "Food desert")

data$Urban <- as.factor(data$Urban)
levels(data$Urban) <- c("Rural", "Urban")
ggplot(data = data,
       aes(x = interaction(LILATracts_1And10, Urban), y = MedianFamilyIncome)) +
      geom_boxplot(fill='white', color="darkred") +
      labs(title="Median Family Income between in urban and rural food deserts",
      x ="Types of tracts ", y = "MedianFamilyIncome") +
      theme(plot.title = element_text(hjust = 0.5))

```

**Urban**: As expected Non-food deserts in urban have higher median income than others.

**Rural**:Contrary to common perceptions, rural food deserts appear to have median family income similar to the urab food deserts.This suggest that the food desert in the rural areas is not solely due to economic reasons but could be influenced by other factors like distribution logistics, infrastructure, or availability of food retailers.

**Non-Food Deserts**: Whether in urban or rural settings, areas that are not classified as food deserts appear to have a more balanced mix of family incomes, suggesting that economic disparities might be less pronounced in these areas.
